---
title: "COGS137 Final Project: California Automobile Accidents 2021"
author: "Adya Mishra, Connor McManigal, Donggyu(Alex) Yu, Mohamed Abdilahi - Data Dream Team"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## Introduction
- INTRODUCTION HERE

### Load packages
```{r load-packages, message=FALSE}
library(tidyverse)
library(tidymodels)
library(olsrr)
library(skimr)
library(lubridate)
```

## Questions

**1. In 2021, is there a relationship between the severity of car accidents and weather conditions in California?**

**2. In San Diego County, is there a relationship between severity of accidents and road infrastructure?**

## The Data
- TALK ABOUT THE DATA HERE
- Explain variables(both ones we ended up using and not using)
- Original data set: 2,845,342 observations and 47 variables
- Wrangled CA data set: 388,838 observations and 37 variables
- SD data set: 24,177 observations and 37 variables

### Data Import
```{r, message = FALSE, warning = FALSE}
crash <- read_csv("/home/cmcmanig/COGS 137/data_dream_team/US_Accidents_Dec21_updated.csv")
```

### Data Wrangling

#### Extract Year and Filter For California Accidents That Happened in 2021
- 
```{r, eval = FALSE}
crash$Year <- substr(crash$Start_Time, 1, 4)
crash2021 <- crash[crash$Year == "2021", ]

crashCA <- crash2021 |>
  filter(State == "CA")
```

#### Extract Month and Day: Create Month and Day Variables
- 
```{r, eval = FALSE}
crashCA$Month <- substr(crashCA$Start_Time, 6, 7)
```

```{r, eval = FALSE}
crashCA$Day <- substr(crashCA$Start_Time, 9, 10)
```

#### Select Desired Variables and Arrange In An Order That Makes Sense
- 
```{r, eval = FALSE}
crashCA <- crashCA |>
  select("ID", "Day", "Month", "Year", "Zipcode", "County", "City", "Street", "Severity", "Start_Lat", "Start_Lng", "Distance(mi)", "Side", "Sunrise_Sunset", "Temperature(F)", "Wind_Chill(F)", "Humidity(%)", "Pressure(in)", "Visibility(mi)", "Wind_Direction", "Wind_Speed(mph)", "Precipitation(in)", "Weather_Condition", "Amenity", "Bump", "Crossing", "Give_Way", "Junction", "No_Exit", "Railway", "Roundabout", "Station", "Stop", "Traffic_Calming", "Traffic_Signal", "Turning_Loop")

head(crashCA)
```

#### Rename Variables Containing Units & Change Variables to Only First Letter Capitalized 
- 
```{r, eval = FALSE}
crashCA <- crashCA |>
  rename(Latitude = "Start_Lat",
         Longitude = "Start_Lng",
         Distance = "Distance(mi)",
         Temperature = "Temperature(F)",
         Wind_chill = "Wind_Chill(F)",
         Humidity = "Humidity(%)",
         Pressure = "Pressure(in)",
         Visibility = "Visibility(mi)",
         Wind_direction = "Wind_Direction",
         Wind_speed = "Wind_Speed(mph)",
         Precipitation = "Precipitation(in)",
         Weather_condition = "Weather_Condition",
         Give_way = "Give_Way",
         No_exit = "No_Exit",
         Traffic_calming = "Traffic_Calming",
         Traffic_signal = "Traffic_Signal",
         Turning_loop = "Turning_Loop",
         Time_of_day = "Sunrise_Sunset")
```

#### Save and Load Wrangled Data Because of Memory in RStudio
- 
```{r, eval = FALSE, echo = FALSE}
save(crashCA, file = "wrangled_crashCA.rda")
```

```{r, eval = FALSE, echo = FALSE}
load("wrangled_crashCA.rda")
```

#### Arrange in Chronological Order and Change Variable Types
- 
```{r, eval = FALSE}
crashCA <- crashCA |>
  arrange(Month, Day) |>
  mutate(Day = as.numeric(Day),
         Month = as.numeric(Month),
         Year = as.numeric(Year))
```

#### Create Date Variable
-
```{r, eval = FALSE}
crashCA$Date <- dmy(paste(crashCA$Day, crashCA$Month, crashCA$Year, sep = "-"))
```

#### Rearange Variables
- 
```{r, eval = FALSE}
crashCA <- crashCA |>
  select("ID", "Day", "Month", "Year", "Date", "Zipcode", "County", "City", "Street", "Severity", "Latitude", "Longitude", "Distance", "Side", "Time_of_day", "Temperature", "Wind_chill", "Humidity", "Pressure", "Visibility", "Wind_direction", "Wind_speed", "Precipitation", "Weather_condition", "Amenity", "Bump", "Crossing", "Give_way", "Junction", "No_exit", "Railway", "Roundabout", "Station", "Stop", "Traffic_calming", "Traffic_signal", "Turning_loop")
```

#### Filter For San Diego County
- 
```{r, eval = FALSE}
crashSD <- crashCA |>
  filter(County == "San Diego")
```

#### Save the Data
- 
```{r, eval = FALSE}
save(crashCA, crashSD, file = "wrangled_crashCA_crashSD.rda")
```


## Analysis
```{r}
load("wrangled_crashCA_crashSD.rda")
```

### Exploratory Data Analysis

```{r}
skim(crashCA)
```

```{r}
skim(crashSD)
```



### Data Analysis

#### Question 1: In 2021, is there a relationship between the severity of car accidents and weather conditions in California?



#### Q1: What's the answer?


#### Question 2: In San Diego County, is there a relationship between severity of accidents and road infrastructure and conditions?



#### Q2: What's the answer?



### Results
- RESULTS HERE

## Conclusion
- CONCLUSION HERE
